---
title: "02-date-exporation"
output: html_document
---
## Load the data
```{r}
library(data.table)
library(dplyr)
x = fread("data-sampled/train_date.csv") %>% as.data.frame()
response = fread("data-sampled/train_numeric.csv", select=c("Response"))$Response %>% factor()
id = x$Id
x$Id <- NULL
dim(x)
head(colnames(x))
tail(colnames(x))
```

## Inspection of the columns reveals them to be numerical with many NA values
Take a few columns at random
```{r}
summary(x[,2])
summary(x[, 100])
summary(x[, 700])
summary(x[, 999])
```

## Make a plot showing the number of NAs
```{r}
pctNA <- function(x){
  a = is.na(x)
  b = sum(a)
  c = length(x)
  d = (b/c)*100
  return(d)
}
percent_na = apply(x, 2, pctNA)
plot(percent_na)
```

## Some interesting patterns
```{r}
column_names = colnames(x)
head(column_names)
```

The Kaggle guidance gives the column description as
aa_bb_dd
aa = Production line
bb = feature
dd = day?

## Production line
```{r}
splits = strsplit(column_names, "_", fixed = T)
production_line = lapply(splits, function(x) x[[1]]) %>% as.character() %>%  factor()
summary(production_line)
plot(x = production_line, y = percent_na, xlab = "Production line", ylab = "Percent of NA values")
```

Production line L0 - L2 are sparse while L3 is not
## Production line feature
```{r}
feature = lapply(splits, function(x) x[[2]]) %>% as.character() %>%  factor()
summary(feature)
table(feature, production_line)
```

Features are specific to a production line

## Third part of string
```{r}
d = lapply(splits, function(x) x[[3]]) %>% as.character() %>%  factor()
length(levels(d)) == length(column_names)
```
Just a column identifier

## Can a Item go accross several production lines?
```{r}
productionLineIndicator <- function(x, line){
  #Function to apply to each row
  f = production_line == line
  a = sum(!is.na(x[f]))
  b = ifelse(a == 0, F, T)
  return(b)
}
L0 = apply(x, 1, productionLineIndicator, line = "L0")
L1 = apply(x, 1, productionLineIndicator, line = "L1")
L2 = apply(x, 1, productionLineIndicator, line = "L2")
L3 = apply(x, 1, productionLineIndicator, line = "L3")


test = L0 + L1 + L2 + L3
summary(test)
hist(test, main = "Number of production lines for each ID")
```

Yes

## Failure rate per production line
```{r}
response = as.character(response) %>% as.numeric()
means = numeric()
sds = numeric()
for(line in levels(production_line)){
  a = mean(response[get(line)])
  b = sd(response[get(line)])
  means = c(means, a)
  sds = c(sds, b)
}

plot_data = data.frame(line = factor(levels(production_line)),
                       mean_response = means,
                       ymax = means + sds,
                       ymin = means - sds)
library(ggplot2)
ggplot(plot_data, aes(x = line, y = mean_response)) +
    geom_point(size = 3) 


```

## With error bars of +- 1sd
```{r}
ggplot(plot_data, aes(x = line, y = mean_response)) +
    geom_point(size = 2) +
    geom_errorbar(aes(ymin = ymin, ymax = ymax))
```

Possibly not significant


## Date encoding
Kaggle page says they are a timestamp
```{r}
all_dates = as.matrix(x) %>% as.vector()
all_dates = all_dates[!is.na(all_dates)]
hist(all_dates, breaks = 150)
```

